---
title: "a3_task4_madeleine_frick"
author: "MaddieFrick"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
```

Read in the urchins Data

```{r}
urchins <- read.csv(here("data","urchins.csv")) %>% #read in the urchins data
  clean_names() #clean the names up
```

Wrangle the urchins data to just include Naples Reef observations for purple urchins observed in Jan 2010

```{r}
purchinjan2010 <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% #filter to only purple urchins
  mutate(date = mdy(date)) %>% #convert the date to class date
  mutate(month = month(date)) %>% #add a new only month column
  mutate(year = year(date)) %>% #add a year column
  filter(month == "1") %>% #only observations in the first month (January)
  filter(year == "2010") %>% #only in the year 2020
  filter(site == "NAPL") #only for Naples site
```

Look and consider histograms and qqplots for the size distribution

```{r}
ggplot(data = purchinjan2010, aes(x = size)) + #histogram with the data as purple urchin one
  geom_histogram(stat = "bin") + #make a histogram
  facet_wrap(~treatment) #make two histograms according to the two different treatments

ggplot(data = purchinjan2010, aes(sample = size)) +
  geom_qq()+ #make a qqplot
  facet_wrap(~treatment) #make the qqplot different according to treatment
```

The data for the annual urchins is in a comb pattern with a mostly uniform spread and the control urchins looks more normally distributed with a skew to the right. It looks like the annual purple urchin size in Naples Reef in January 2010 might be larger for the annual than the control group. The qqplot shows us that the data for both different treatments is not necessarily very close to normal, as they do not produce straight lines when the theoretical is graphed against the sample. The lines have more of a curve to them, showing them that there is not very much distribution.

```{r}
purchin_summary <- purchinjan2010 %>% #take the purple urchin dataset
  group_by(treatment) %>% #group by the different treatments
  summarize(mean = mean(size),#finds the mean of the size measurements
            median = median(size), #same but ^
            sd = sd(size), #again
            n = n()) %>% 
  kable(col.names = c("Treatment", #this names all of the columns
                      "Mean",
                      "Median", 
                      "Standard Deviation",
                      "Sample Size")) %>% 
  kable_styling() #make the table pretty

purchin_summary #show the table

```


```{r}

annual_sample <- purchinjan2010 %>% #this is making a vector of the annual sizes
  filter(treatment == "ANNUAL") %>% 
  pull(size)

control_sample <- purchinjan2010 %>% #this is making a vector of the control sizes
  filter(treatment == "CONTROL") %>% 
  pull(size)
#then run a two sample t-test on these vectors

t.test(annual_sample, control_sample) #run a t test on x, being the annual and y being the control

```

Since the p value is somewhat relatively high, there is less evidence of differences between the two population means of the control and annual samples. The 95% confidence interval of -0.13 to 0.52 means that we are 95% confident that the difference between the two means is between that amount of each other, where 0 is included, 0 being no difference in the means. The actual mean of the annual group was 3.9 and the actual mean of the control group was 3.7, which are close to the same size. The standard deviation for the annual group was .96 and .86 for the control group, both showing around 1. This means if they were plotted on the same graph with the mean value near 3.8 in the center, one standard deviation of about 1 on either side of both groups would mainly overlap one another. The meaning of tje p value is that there is a 25.3% chance that I found two samples with means that are least as different as the ones that I found by random chance if the null hypothesis was true; they were drawn from populations with the same mean.
